Reproducible Research Assignment 1
========================================================


```{r}
install.packages("data.table")
library(data.table)

install.packages("ggplot2")
library(ggplot2)


```
### 1.  Load the data (i.e. read.csv())

```{r}
activityDF <- read.csv("activity.csv" , sep="," , header =TRUE)
```

### What is mean total number of steps taken per day?
### For this part of the assignment, you can ignore the 
### missing values in the dataset.

```{r}
newdata <- subset(activityDF, steps != "NA", select=c(steps, date ,interval))
newdataDT <- data.table(newdata)
newdataFinal <- newdataDT[,sum(steps) , by = date]
plotdata <- as.data.frame(newdataFinal)

```

### 1.  Make a histogram of the total number of steps taken each day

```{r fig.width=7, fig.height=6}
hist(plotdata$V1 , main= "Activity Monitoring Data",col= "red",  xlab="Steps")
```


### 2.  Calculate and report the mean and median total number 
### of steps taken per day


```{r}
mean(plotdata$V1)
median(plotdata$V1)

```

### What is the average daily activity pattern?

```{r}
newdata1 <- subset(activityDF, steps != "NA" , select=c(steps, date ,interval))
DT <- data.table(newdata1)
DTfinal <- DT[, sum(steps) , by = interval ]

```
### 1.  Make a time series plot (i.e. type = "l") of the 
### 5-minute interval (x-axis) and the average number of steps taken,
### averaged across all days (y-axis)

```{r fig.width=7, fig.height=6}
qplot(interval, V1, data = as.data.frame(DTfinal), geom = "line" ,main="Activity" , ylab="Steps")

```


### 2.  Which 5-minute interval, on average across all the days 
### in the dataset, contains the maximum number of steps?


```{r}
tail(DTfinal[order(DTfinal$V1  ), ], n=1)

```

### Imputing missing values
### Note that there are a number of days/intervals where there are 
### missing values (coded as NA). The presence of missing days may 
### introduce bias into some calculations or summaries of the data.

### 1.  Calculate and report the total number of missing values in 
### the dataset (i.e. the total number of rows with NAs)


```{r}
sum(is.na(activityDF)) 
activityDF1 <- activityDF
DT3 <- data.table(activityDF1)

```

### 2.  Devise a strategy for filling in all of the missing values 
### in the dataset. The strategy does not need to be sophisticated. 
### For example, you could use the mean/median for that day, 
### or the mean for that 5-minute interval, etc.


```{r}
DTMean <- DT[, mean(steps) , by = interval ]

MrgDT = merge(DT3 , DTMean , by ="interval" , suffixes =c(".DT3" , ".DTMean"))
na.idx = which(is.na(DT3$steps))
DT3[na.idx , "steps"] = MrgDT[na.idx ,"V1.DTMean"]

```
### 3.  Create a new dataset that is equal to the original 
### dataset but with the missing data filled in.


```{r}

MrgDT$steps[is.na(MrgDT$steps) ] <- MrgDT$V1

MrgDT$steps <- round(MrgDT$steps ,2)

MrgDTFinal <- MrgDT[,sum(steps) , by = date]

plotdata3 <- as.data.frame(MrgDTFinal)
```


### 4.  Make a histogram of the total number of steps taken 
### each day and Calculate and report the mean and median total 
### number of steps taken per day. Do these values differ from the 
### estimates from the first part of the assignment? What is the 
### impact of imputing missing data on the estimates of the total 
### daily number of steps?

```{r fig.width=7, fig.height=6}

hist(plotdata3$V1 , main= "Activity Monitoring Data",col= "orange",  xlab="Steps")

```


```{r}
mean(plotdata3$V1)

median(plotdata3$V1)


```

### Are there differences in activity patterns between weekdays and weekends?
### For this part the weekdays() function may be of 
### some help here. Use the dataset with the filled-in 
### missing values for this part.


```{r}
MrgDF <- as.data.frame(MrgDT)

MrgDF[,"date"] <- as.Date(as.character(MrgDF[,"date"]))

MrgDF$Wday <- weekdays(MrgDF[ , "date"] , abbreviate = TRUE)


```

### 1.  Create a new factor variable in the dataset 
### with two levels - "weekday" and "weekend" indicating 
### whether a given date is a weekday or weekend day.


```{r}
MrgDF$Wday[MrgDF$Wday %in% c("Mon" ,"Tue" ,"Wed" ,"Thu" ,"Fri")] <- "Weekday"

MrgDF$Wday[MrgDF$Wday %in% c("Sat" ,"Sun")] <- "Weekend"

MrgDT1 <- data.table(MrgDF)


```


### 2.  Make a panel plot containing a time series plot 
### (i.e. type = "l") of the 5-minute interval (x-axis) and 
### the average number of steps taken, averaged across all 
### weekday days or weekend days (y-axis). 

```{r}
setkey(MrgDT1 , interval , Wday)

MrgDT1final <-  MrgDT1[,sum(steps) , by = key(MrgDT1)]


```

```{r fig.width=10, fig.height=6}


qplot(x = interval ,y = round(V1) , data = as.data.frame(MrgDT1final) , group =1)+ facet_grid(Wday ~ . )  + geom_line() +xlab("Interval") +ylab("Steps") + ggtitle("Activity monitoring data")


```



